Broom: Safe Region-Based Memory Management for Dataflow Programs

### 2015-11-09 paper/introduction.tex
updates to introduction

\subsection*{Outline}

@@

The goal of this work is an automatic (memory-safe) region-based memory management technique
that supports dynamic regions and higher-order functions.
Though our focus is on memory-safety, dynamic regions also have other side-benefits.
When a producer transfers a data-structure to a consumer, in a shared-memory context , our approach
provides an ownership discipline that ensures that the producer cannot subsequently modify
the data-structure. When the same transfer happens in a distributed setting, the system guarantees
that the data-structure is self-contained and can be safely transmitted across address spaces.
\TODO{How accurate is this claim?}

@@

\begin{itemize}

\item
This reduces the problem of ensuring memory safety to two sub-problems: the first one, to ensure the inter-region pointer invariants about which objects in which regions can point to which objects in which regions, and the second one, ensuring the correct usage protocol for pointers (handles) to regions themselves. In particular, the second one includes the constraints such as: do not free an open region, do not open a freed region, etc.

\item
We can elaborate on these two categories on invariants.

\item
We then explain that our type-system statically ensures that the program will satisfy the first set of constraints.

\item
We explain that we verify that the second set of constraints are satisfied dynamically.

\item
We justify our choice to do in this fashion: the first set of constraints are expensive to check dynamically, while the second set of constraints is not. In contrast, a static typechecker that rules out the second category of violations would be very restrict from an expressiveness perspective (requiring a linear type system). Finally, the human effort to ensure the second set of invariants is not as burdensome (as these are coarse-grained objects manipulated infrequently), unlike in the case for the first set.

\item
We can perhaps illustrate the need to create handles/pointers to regions and store them in dynamic data-structures like dictionaries to illustrate why restricting the usage of such pointers would be too constraining.

>>>>

@@

The goal of this work is an automatic (memory-safe) region-based memory management technique
that supports dynamic regions (that are first-class objects) and higher-order functions.
Though our focus is on memory-safety, dynamic regions also have other side-benefits.
When a producer transfers a data-structure to a consumer, in a shared-memory context , our approach
provides an ownership discipline that ensures that the producer cannot subsequently modify
the data-structure. When the same transfer happens in a distributed setting, the system guarantees
that the data-structure is self-contained and can be safely transmitted across address spaces.
\TODO{How accurate is this claim?}

@@

The final piece required to ensure memory-safety is ensuring that the program correctly follows the above-mentioned
protocol for dynamic regions. The protocol is essentially a typestate specification: an open region cannot be freed and,
dually, a freed region cannot be opened.

Dynamic regions are first-class objects in our system: e.g., in our running example,
we would like to use a dictionary that stores dynamic regions. However, this means that the program may create multiple
aliases (pointers) to the same region, which, in turn, means that statically checking if the program correctly follows the
dynamic region protocol is hard (undecidable, in fact). Safety can be ensured by using, \eg, linear typing to prevent the
creation of aliases, but this would be quite restrictive. Hence, in our system, we dynamically check for adherence to
this protocol.

In summary, our approach reduces memory safety to two sub-problems:
(a) Ensuring  low-level invariants about pointers to (fine-grained) objects inside regions and 
(b) Ensuring higher-level (typestate) invariants about the regions themselves. 
We use a type system to statically ensure the first set of invariants. We use a dynamic check to ensure that the
second set of invariants are satisfied.
%
This is a novel aspect of our system. We believe that this is a reasonable choice:
dynamically checking for the second category of invariants is easy and does not pose a performance concern (unlike, \eg, the first category of constraints);
a static typechecker that rules out the second category of violations would be very restrictive from an expressiveness perspective;
finally, the human effort required to ensure the second set of invariants is not as burdensome (as regions are coarse-grained objects
manipulated infrequently), \eg, as it would be for the first set of invariants.
